---
title: "Predicting Wine Quality with Neural Networks"
subtitle: "Neural Networks"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "Havva Nur Elveren"
output:
  html_document:
      theme: journal
      toc: yes
      toc_depth: 4
      #toc_float: true
  word_document:
      toc: yes
      toc_depth: 4
      #toc_float: true
  pdf_document:
      toc: yes
      theme: journal
      toc_depth: 4
      #toc_float: true
---
---
# Objective: Predicting Wine Quality
Can we predict wine quality based on its features such as acidity, alcohol, sugar or sulfate level? In this project, we'll predict Wine Quality with looking at the value of different features of a wine. We'll use a data set that has been collected from red wine variants of the Portuguese "Vinho Verde" wine. If quality is greater than 6.5 it is considered as good wine, otherwise it is considered as bad wine.

# Data Description:
* 1.6K Row with 12 Column. You can download the data from the link https://www.kaggle.com/uciml/red-wine-quality-cortez-et-al-2009
```{r}
library(kableExtra)

dt <- data.frame(Name = c("fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide",
"density", "pH", "sulphates", "alcohol", "quality"),
Description = c("most acids involved with wine or fixed or nonvolatile (do not evaporate readily)", "the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste", "found in small quantities, citric acid can add 'freshness' and flavor to wines", "the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than", "the amount of salt in the wine", "the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents", "amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2", "the density of water is close to that of water depending on the percent alcohol and sugar content", "describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the", "a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and", "the percent alcohol content of the wine","Score between 0 and 10, if quality > 6.5 it's Good, otherwise it is Bad "))

dt %>%
  kbl() %>%
  kable_styling()

```
## STEP 1: Load the Libraries

```{r message=FALSE, warning=FALSE}
library(ISLR)
library(caret)
library(readxl)
library(pROC)
library(lattice)
library(ggplot2)
library(dplyr)
library(e1071) 
library(corrplot)
library(kknn)
library(ggplot2)
library(multiROC)
library(MLeval)
library(AppliedPredictiveModeling)
library(corrplot)
library(Hmisc)
library(dplyr)
library(quantmod) 

library(nnet)
library(caret)
library(NeuralNetTools)
library(gamlss.add)
library(neuralnet)

```

## Step 2: Load the Data Set
```{r message=FALSE, warning=FALSE}
winedata <- read.csv("winequality-red.csv")
winedata <- data.frame(winedata, stringsAsFactors = FALSE, colnames<- TRUE )

str(winedata)
summary(winedata)

# If quality score is greater than 6.5 set quality as Good, otherwise set as Bad
winedata$quality[winedata$quality>6.5] <- 'Good'
winedata$quality[winedata$quality<= 6.5] <- 'Bad'
winedata$quality <- as.factor(winedata$quality)
table(winedata$quality)

```

## Step 3: Prepare the Test and Training Data
```{r message=FALSE, warning=FALSE}
# Use the %75 of the data for training and the rest for testing the model
trainingIndex <- createDataPartition(y = winedata$quality, p = .75, list = FALSE)
winedata_train <- winedata[trainingIndex,]
winedata_test <- winedata[-trainingIndex,]

# with excluding the quality column, scale the data with z-score
trainX <- winedata_train[,names(winedata_train) != "quality"]
preProcValues <- preProcess(x = trainX,method = c("center", "scale"))
preProcValues

```

## Step 4: Train the Model with Neural Network Algorithm
```{r message=FALSE, warning=FALSE}
set.seed(123)
# Use K-fold cross validation with k=10 for training to split the training data randomly to k subset and train the model repeatedly based on given repeat value.
control <- trainControl(method = "repeatedcv", number = 10, repeats = 1)

# Train the model with Neural Network algorithm using nnet function.
model <- train(quality~ ., data = winedata_train, method="nnet", trControl= control, preProcess=c("scale","center"), na.action = na.omit, trace=FALSE, verbose=FALSE)

model$results
plotnet(model, y_names = "Wine Quality")

# Make prediction with using Neural Network model
prediction <- predict(model, winedata_test[-12])                           
table(prediction, winedata_test$quality)   

confusionMatrix(prediction, winedata_test$quality)

```

## Step 5: Improve Model Performance 
```{r message=FALSE, warning=FALSE}
# Improve the model performance by increasing the number of layers of the neural network and only using the important predictors.

#Plot Variable Importance 
varImp<-varImp(model)
print(varImp)
ggplot(varImp)
plot(varImp)

#Increase the number of hidden layers in the model.
nnetGrid <-  expand.grid(size = seq(from = 1, to = 10, by = 1),
                        decay = seq(from = 0.1, to = 0.5, by = 0.1))

# Train the model with the most important predictors.
model <- train(quality~ sulphates + total.sulfur.dioxide + citric.acid + alcohol + fixed.acidity + residual.sugar, data = winedata_train, method="nnet", trControl= control, preProcess=c("scale","center"), na.action = na.omit, tuneGrid = nnetGrid, trace=FALSE, verbose=FALSE)

# Plot the model and print the model results.
model$results
plotnet(model, y_names = "Wine Quality")

# Make prediction with using Neural Network model
prediction <- predict(model, winedata_test[-12])                           
table(prediction, winedata_test$quality)   

confusionMatrix(prediction, winedata_test$quality)

```

# Conculusion

In this project we predict wine quality with using neural network algorithm. To train the model we use "nnet" function which uses logistic sigmoid function for hidden layer activation with default one hidden layer. 

Before sarting to train our model, first we load the data and assign good and bad labels to each wine data according to it's quality score. If the quality is grater than 6.5 than it's considered as good wine, otherwise it's considered as bad wine.

After preparing the test and training data, we train our model with nnet function. We use k-fold cross validation for random sampling. 

Our first model which includes all the features as predictor, predicts the wine quality with %88 accuracy. It's a good accuracy rate we can confidently use this model to predict wine quality. 
After creating our model with default nnet function, we try to improve model performance by increasing the hiddlen layers in the neural network and only using features that has more than %50 importance rate. And we can see that this new model predicts the wine quality with slightly higher accuracy. New model doesn't provide a drastic performance improvement but still have a lower error rate and better accuracy than our first model.

Both model predicts the bad quality wines better than good quality wines as expected. Because our data is not balanced and has more bad wine data than good wine data.

Looking at the results of both models we can confidently say that neural network algorithm is a good choice to make prediction on our wine data set.
